### 2021-06-18 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Added comments/questions
minor edits after readthrough (typos, repetition of accronym defs, etc.)

For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results. In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}. The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
\bseqn{
	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
}
where the radial part is
\bseqn{
	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
}
for $n - m$ even, while it is defined as $0$ for $n - m$ odd. A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}. These radial OPs are:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$. These OPs constitute a one-parameter family of orthogonal polynomials on the unit disk, with the classical Zernike polynomials being retrieved via setting the parameter $k = 0$. The use of polar coordinates allows for radially symmetric partial differential operators, while the incorporation of a parameter means they can be represented as banded matrices by exploiting the relations of the Jacobi polynomials. Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

@@

We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}. On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
\bseqn{
	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c
}
for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$. The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
\bseqn{
	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
}
where $C_{n,k} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.

@@

The comparison with the construction of the 3D spherical harmonics is evident here (the $\genjac$ OPs are in place of the Jacobi polynomials, and the $\jac$ OPs are in place of the complex exponentials, which are just Chebyshev polynomials). By defining our OPs in this way, we will be able to yield sparse and banded relations for the necessary operators involved in PDEs, just as we did before.

@@

Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method. To obtain the entries for them, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$. These can be expressed as: 
\bseqnnumber{
	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:ds:Hrecurrence} \\
	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
}
Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$. We can use \bsrefeqns{eqn:ds:Hrecurrence}{eqn:ds:Precurrence} to determine the 2D recurrences for $\hdopnkabcd(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships.

\begin{lemma}
	$\hdopnkabcd(x,y)$ satisfy the following 3-term recurrences:
\bseqn{
	x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
	y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
}
for $(x,y) \in \Omega$, where
\bseqn{
	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)} \\
	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}. 
}

@@

\begin{proof}
The 3-term recurrence for multiplication by $x$ follows from \bsrefeqn{eqn:ds:Hrecurrence}. For the recurrence for multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$. These coefficients are given by \nomenclature[abs]{$\abs{m}$}{The absolute value of a scalar $m$} \nomenclature[norm]{$\norm{f}$}{The norm of a function f, defined by the square root of the relevant inner product} \nomenclature[normH]{$\norm{f}_{\hdopabcd}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
\bseqn{
	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\hdopabcd}}{\norm{\hdopmjabcd}^{-2}_{\hdopabcd}}.
}
Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopmjabcd}_{\hdopabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$. Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
\bseqn{
	&\ip<y \hdopnkabcd, \hdopmjabcd>_{\hdopabcd} \\
	&=  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b,c+d)}(x) \: \D x \Big) \\
		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \genjacw^{(a, b,c+d+k+j+2)}(x) \: \D x \Big) \\
		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
	&=
	\begin{cases}
    		\delta_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+3)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k+1 \\
		\gamma_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+1)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k \\
		\delta_{k-1}^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k-1)} \: \ip<\genjacnmk^{(a, b, c+d+2k-1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k-1 \\
		\quadeight \quadfour 0 \quadfour \quadeight \text{otherwise}
      	\end{cases}
}
where, by orthogonality,
\bseqn{
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} &= 0 \qfor m < n-1, \\
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} &= 0 \qfor m < n-1.
}

@@

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: \nomenclature[Hvectorabcd]{$\bighdopabcd$}{Vector of the disk-slice/trapezium OPs ordered by degree, with parameters $a,b,c,d$} \nomenclature[Hvectornabcd]{$\bighdopabcd_n$}{(Sub-)Vector of the disk-slice/trapezium OPs of degree $n$, with parameters $a,b,c,d$}
\bseqn{
	\bighdopabcd_n := 
		\begin{pmatrix}
			\hdopabcd_{n,0}(x,y) \\
			\vdots \\
			\hdopabcd_{n,n}(x,y)
		\end{pmatrix} \in \R^{n+1}, 
	\quad \quad 
	\bighdopabcd := 
		\begin{pmatrix}
			\bighdopabcd_0 \\
			\bighdopabcd_1 \\
			\bighdopabcd_2 \\
			\vdots \\
		\end{pmatrix}
}
and set $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ as the Jacobi matrices corresponding to
\bseqnnumber{
	J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) &= x \: \bighdopabcd(x,y), \label{eqn:ds:jacobimatricesdefinitionx} \\
	J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) &= y \: \bighdopabcd(x,y). \label{eqn:ds:jacobimatricesdefinitiony}
}
The matrices $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkabcd\}$ basis. For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \vec{f}$. Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b,c,d)\top}} \vec{f}$. In other words, ${J_x^{(a,b,c,d)\top}} \vec{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the $\{\hdopnkabcd\}$ basis.

@@

As eluded to, in general we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example. However, here we simply consider our domain to be a disk-slice and our discretisation to be a single element -- that is we let $\element = \Omega$ for a disk-slice domain. We can choose our finite dimensional space $V_p = \bsset{v_p \in V}{\deg(v_p|_\element) \le p}$ for some $p \in \N$.

We seek $u_p \in V_p$ s.t.
\bseqnnumber{
	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
	\label{eqn:ds:FEMweakform}
}

>>>>

A review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results. In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}. The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
\bseqn{
	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
}
where the radial part is
\bseqn{
	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
}
for $n - m$ even, while it is defined as $0$ for $n - m$ odd. A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}. These radial OPs are:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$. These OPs constitute a one-parameter family of orthogonal polynomials on the unit disk, with the classical Zernike polynomials being retrieved via setting the parameter $k = 0$. The use of polar coordinates allows for radially symmetric partial differential operators, while the incorporation of a parameter means they can be represented as banded matrices by exploiting the relations of the Jacobi polynomials. Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

@@

We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}. On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
\bseqn{
	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c
}
for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$. The corresponding OPs are denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
\bseqn{
	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
}
where $C_{n,k} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.

@@

The comparison with the construction of the 3D spherical harmonics is evident here (the $\genjac_n$ OPs are in place of the Jacobi polynomials, and the $\jac_n$ OPs are in place of the complex exponentials, which are just Chebyshev polynomials). By defining our OPs in this way, we will be able to yield sparse and banded relations for the necessary operators involved in PDEs, just as we did before.

@@

Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method. To obtain their entries, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$. These can be expressed as: 
\bseqnnumber{
	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:ds:Hrecurrence} \\
	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
}
Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$. We can use \bsrefeqns{eqn:ds:Hrecurrence}{eqn:ds:Precurrence} to determine the 2D recurrences for $\hdopnkabcd(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships.

\begin{lemma}
	$\hdopnkabcd(x,y)$ satisfy the following relations:
\bseqn{
	x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
	y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) \\
		&\quad + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
}
for $(x,y) \in \Omega$, where
\bseqn{
	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)} \\
	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}. 
}

@@

\begin{proof}
The relation for multiplication by $x$ follows from \bsrefeqn{eqn:ds:Hrecurrence}. For multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$. These coefficients are given by \nomenclature[abs]{$\abs{m}$}{The absolute value of a scalar $m$} \nomenclature[norm]{$\norm{f}$}{The norm of a function f, defined by the square root of the relevant inner product} \nomenclature[normH]{$\norm{f}_{\hdopabcd}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
\bseqn{
	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\hdopabcd}}{\norm{\hdopmjabcd}^{-2}_{\hdopabcd}}.
}
Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopmjabcd}_{\hdopabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$. Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
\bseqn{
	&\ip<y \hdopnkabcd, \hdopmjabcd>_{\hdopabcd} \\
	&=  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b,c+d)}(x) \: \D x \Big) \\
		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \genjacw^{(a, b,c+d+k+j+2)}(x) \: \D x \Big) \\
		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
	&=
	\begin{cases}
    		\delta_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+3)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k+1 \\
		\gamma_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+1)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k \\
		\delta_{k-1}^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k-1)} \: \ip<\genjacnmk^{(a, b, c+d+2k-1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} \\
			\quadeight \quadeight \quadeight \: \: \text{if } j = k-1 \\
		\quadeight \quadfour 0 \quadfour \quadeight \text{otherwise}
      	\end{cases}
}
where, by orthogonality,
\bseqn{
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} &= 0 \qfor m < n-1, \\
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} &= 0 \qfor m < n-1.
}

@@

These relations lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: \nomenclature[Hvectorabcd]{$\bighdopabcd$}{Vector of the disk-slice/trapezium OPs ordered by degree, with parameters $a,b,c,d$} \nomenclature[Hvectornabcd]{$\bighdopabcd_n$}{(Sub-)Vector of the disk-slice/trapezium OPs of degree $n$, with parameters $a,b,c,d$}
\bseqn{
	\bighdopabcd_n := 
		\begin{pmatrix}
			\hdopabcd_{n,0}(x,y) \\
			\vdots \\
			\hdopabcd_{n,n}(x,y)
		\end{pmatrix} \in \R^{n+1}, 
	\quad \quad 
	\bighdopabcd := 
		\begin{pmatrix}
			\bighdopabcd_0 \\
			\bighdopabcd_1 \\
			\bighdopabcd_2 \\
			\vdots \\
		\end{pmatrix}
}
and set $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ as the Jacobi matrices corresponding to
\bseqnnumber{
	J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) &= x \: \bighdopabcd(x,y), \label{eqn:ds:jacobimatricesdefinitionx} \\
	J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) &= y \: \bighdopabcd(x,y). \label{eqn:ds:jacobimatricesdefinitiony}
}
The matrices $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkabcd\}$ basis. For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \vec{f}$. Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b,c,d)\top}} \vec{f}$. In other words, ${J_x^{(a,b,c,d)\top}} \vec{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the $\{\hdopnkabcd\}$ basis.

@@

As eluded to, in general we would discretise our domain $\Omega$ with a mesh $\FEset$ consisting of elements $\element$, where each $\element \in \FEset$ is a trapezium or disk slice for example. However, here we simply consider our domain to be a disk-slice and our discretisation to be a single element -- that is we let $\element = \Omega$ for a disk-slice domain. We can choose our finite dimensional space $V_p = \bsset{v_p \in V}{\deg(v_p|_\element) \le p}$ for some $p \in \N$.
